### 2024-11-15 Draft.Rmd
Update Draft.Rmd

---
title: "Analysis of Global Crises and Their Impact on Interational Well-Being"
author: "Group 8"
date: "November 12, 2024"
output: pdf_document
---

List your group members, including their student numbers, here:

-   Aditya Chauhan (169027493)
-   Andy Ferrer Lorenzo (169081356)
-   Oscar Saul (169057882)
-   Taras Zakydalsky (169099616)
-   Dibuhang Rai (169708451)

You **must** be in a group in MyLS in order to see the DropBox used for submission. Even if you're alone, you must join a group by yourself.

You **must** be in a group with people from the same section as you. MyLS does not allow for groups including students from both Data100A and Data100B.

```{r initial_setup, include=FALSE}
# echo = FALSE will set the Rmd to *not* show the R code. Don't change this.
# You may change the default figure width and figure height as you please.
knitr::opts_chunk$set(echo = FALSE, message = FALSE, fig.width = 6)
# Put any libraries that you need to load here.
# DO NOT PUT "install.packages()" IN AN RMD FILE!!!
library(tidyverse)
library(arrow)
library(maps)
library(tinytex)
library(sf)
library(rnaturalearth)
library(corrplot)
library(scales)
library(geosphere)
library(ggridges)
```

# Instructions

You are encouraged to remove this instruction section prior to submission.

It is recommended that you follow the structure of this template. The text is all placeholder - you are free to change any/all wording as you please, but it is very helpful for the grading process if you keep the same structure. Anything in \<<double angle brackets>\> definitely needs to be changed, but you are free to change any/all sentences!

Note that all of the code is *hidden* by default. This file will be graded based on the insights, not the code.

You will only submit the PDF version of this document. To knit to PDF, you'll need to run `install.packages("tinytex")` in the console, followed by `tinytex::install_tinytex()` (DO NOT PUT THESE COMMANDS IN AN RMD FILE!!!). If you encounter errors in "Knit to PDF", you can "knit to html" and then print the html file to PDF using your operating system's PDF view (e.g. Adobe Acrobat). Only standalone PDF files will be accepted by MyLS.

# Purpose

This study aims to

# Abstract

# Introduction

Climate change is something that has been studied. Here's some relevant information about the context of our study.

If needed, this paragraph is more information about the context.

In this report, we are going to explore some aspects climate change and the impact and/or perceptions of it by using exploratory techniques. We'll explore \<<general description of data>\> using \<<general description of techniques>\>.

By the end of this report, we will have shown ...

# Data Description

## \<\<Cyclones Data\>\>

```{r load_data1}
cyclones_data <- read_parquet("cyclones_data.parquet")
world <- ne_countries(scale = "medium", returnclass = "sf")

# ________________Map long and lat to scale________________
# Convert latitude(-90/+90) and longitude(-180/+180) to be within the correct range
cyclones_data <- cyclones_data |>
    rename(long = lon) |> #rename lon to long to match with world map
    mutate(
        lat = ((lat + 90) %% 180) - 90,  
        long = ((long + 180) %% 360) - 180 
    )

# ________________Determine distance of cyclone to equator________________
equator_point <- c(0, 0)
cyclones_data <- cyclones_data |>
  mutate(
    km_distance_from_equator = distGeo(cbind(long, lat), cbind(long, 0)) / 1000
  )

# ________________Determine distance of cyclone to equator________________
equator_point <- c(0, 0)
cyclones_data <- cyclones_data |>
  mutate(
    km_distance_from_equator = distGeo(cbind(long, lat), cbind(long, 0)) / 1000
  )

# ________________Convert long and lat points to countries________________
cyclones_data <- st_as_sf(cyclones_data, coords = c("long", "lat"), crs = 4326, remove = FALSE) |>
  st_join(world["name"], left = TRUE) |>
  rename(country = name)

# ________________Create cyclone_cleaned with specified columns________________
cyclone_cleaned <- cyclones_data |>
    select(country, lat, long, category, km_distance_from_equator)

cyclone_cleaned

```

The data set `cyclone_data` provided by the National Hurricane Center and the Central Pacific Hurricane Center records hurricane entries in the Atlantic and North Pacific basins from 1851 to 2022 and 1949 to 2022, respectively. The data frame includes detailed time stamps, coordinates, and categories, which are vital for analysis.

The preliminary step in supporting the analysis involved converting the coordinates into the Cartesian coordinate system. Next, we used the `rnaturalearth` package and `geolocation` to convert the coordinates into country identifiers and determine each storm's distance in kilometers from the equator. As shown in Figure 1 in the appendix, cyclones are now accurately distributed across the North Atlantic and Northeast Pacific basins with the correct identifiers.

## \<\<World Covid 2020\>\>

```{r load_data2}
covid_2020 <- read_parquet("covid_2020.parquet")
world <- ne_countries(scale = "medium", returnclass = "sf")

covid_2020 <- covid_2020 |>
  mutate(country = case_when(
  country == "British Virgin Islands" ~ "British Virgin Is.",
  country == "Cape Verde" ~ "Cabo Verde",
  country == "Cayman Islands" ~ "Cayman Is.",
  country == "Central African Republic" ~ "Central African Rep.",
  country == "Cook Islands" ~ "Cook Is.",
  country == "Cote d'Ivoire" ~ "Ivory Coast",
  country == "Democratic Republic of Congo" ~ "Congo",
  country == "Cyprus" ~ "Cyprus",
  country == "Dominican Republic" ~ "Dominican Rep.",
  country == "East Timor" ~ "Timor-Leste",
  country == "England" ~ "United Kingdom",
  country == "Equatorial Guinea" ~ "Equat. Guinea",
  country == "Eswatini" ~ "Swaziland",
  country == "Falkland Islands" ~ "Falkland Isles",
  country == "Faroe Islands" ~ "Faroe Isles",
  country == "French Guiana" ~ "Fr. Guiana",
  country == "French Polynesia" ~ "Fr. Polynesia",
  country == "Gibraltar" ~ "Gibraltar",
  country == "Guadeloupe" ~ "Guadeloupe",
  country == "Marshall Islands" ~ "Marshall Is.",
  country == "Martinique" ~ "Martinique",
  country == "Mayotte" ~ "Mayotte",
  country == "Micronesia (country)" ~ "Micronesia",
  country == "Northern Ireland" ~ "Northern Ireland",
  country == "Northern Mariana Islands" ~ "N. Mariana Is.",
  country == "Pitcairn Islands" ~ "Pitcairn",
  country == "Reunion" ~ "Réunion",
  country == "Saint Barthelemy" ~ "Saint-Barthélemy",
  country == "Saint Kitts and Nevis" ~ "St Kitts & Nevis",
  country == "Saint Martin (French part)" ~ "Saint Martin",
  country == "Saint Pierre and Miquelon" ~ "St Pierre & Miquelon",
  country == "Saint Vincent and the Grenadines" ~ "St Vincent & Gren.",
  country == "Sao Tome and Principe" ~ "São Tomé",
  country == "Scotland" ~ "Scotland",
  country == "Sint Maarten (Dutch part)" ~ "Sint Maarten",
  country == "Solomon Islands" ~ "Solomon Is.",
  country == "South Sudan" ~ "S. Sudan",
  country == "Tokelau" ~ "Tokelau",
  country == "Turks and Caicos Islands" ~ "Turks & Caicos",
  country == "United States Virgin Islands" ~ "US Virgin Is.",
  country == "Wales" ~ "Wales",
  country == "Wallis and Futuna" ~ "Wallis & Futuna",
  country == "United States" ~ "United States of America",
  TRUE ~ country  
))

equator <- st_sfc(st_linestring(rbind(c(-180, 0), c(180, 0))), crs = st_crs(world))
world <- st_transform(world, crs = 3857)
equator <- st_transform(equator, crs = 3857)

world_with_distance <- world |>
  mutate(
    centroid = st_centroid(geometry),
    latitude = st_coordinates(centroid)[, 2],
    hemisphere = ifelse(latitude > 0, "North", "South"),
    distance_to_equator_km = as.numeric(st_distance(centroid, equator)) / 1000  # Convert to kilometers
  ) |>
  select(name, hemisphere, distance_to_equator_km) |>
  rename(country = name)

covid_2020 <- covid_2020 |>
  left_join(world_with_distance, by = "country")

covid_2020

```

The `covid_2020` data set, provided by Our World in Data contains a comprehensive list of the total reported cases of COVID-19 cases across many countries and continents in 2020. To aid in the analysis `rnaturalearth` and `geolocation` was used to find the hemisphere and distance to the equator in kilometers.

## \<\<World Happiness Census\>\>

```{r load_data3}
# Put in your code to load in the data set, along with any
# necessary cleaning beyond what was done in Part 1
happiness <- read_parquet("happiness.parquet")
world_map <- map_data("world")

happiness_mapped <- happiness |>
    mutate(country = case_when( #match country names to world_map
        country == "United States" ~ "United States of America",
        country == "United Kingdom" ~ "UK",
        country == "Congo (Brazzaville)" ~ "Congo",
        country == "Congo (Kinshasa)" ~ "Congo",
        country == "Hong Kong S.A.R. of China" ~ "Hong Kong",
        country == "Türkiye" ~ "Turkey",
        country == "Lao.People's.Democratic.Republic" ~ "Laos",
        country == "United Arab Emirates" ~ "UAE",
        TRUE ~ country
    )) |>
    right_join(world_map, by = c("country" = "region")) #joins the long and lat of the respective countries to map them

world <- ne_countries(scale = "medium", returnclass = "sf")

equator <- st_sfc(st_linestring(rbind(c(-180, 0), c(180, 0))), crs = st_crs(world))
world <- st_transform(world, crs = 3857)
equator <- st_transform(equator, crs = 3857)

happiness <- happiness |>
  left_join(
    world |>
      mutate(
        centroid = st_centroid(geometry),
        latitude = st_coordinates(centroid)[, 2],
        hemisphere = ifelse(latitude > 0, "North", "South"),
        distance_to_equator_km = as.numeric(st_distance(centroid, equator)) / 1000
      ) |>
      select(name, hemisphere, distance_to_equator_km) |>
      rename(country = name),
    by = "country"
  )
happiness

```

the `happiness` dataset was loaded, containing data on countries' economic, social, and political stability from the World Happiness Report. To ensure consistency across datasets, some country names were adjusted to match the naming conventions in the `world_map` dataset. This adjustment was necessary for accurate geographic mapping. After renaming, the `happiness` dataset was joined with `world_map` to incorporate latitude and longitude coordinates for each country, enabling the visualization of global data trends on a map. Additional geographic information was also added, including the distance of each country from the equator and hemisphere designation, by using centroids of each country's polygon in the `world` dataset. This enhancement facilitates spatial analysis based on proximity to the equator and hemisphere differences.

## Combining the Data

Explain how any combinations of data were performed. Explain what kind of join was needed, whether columns had to be modified (for example, matching "country" names.)

# Exploratory Data Analysis

We explored many aspects of the data, but will demonstrate three. These are \<\<Geographic and environmental Factors VS a countries overall well being\>\>, \<\<Covid Cases Vs perception of corruption\>\>, and \<\<Freedom to Make Life Choices & Climate Awareness (OR DISTANCE VS COVID)\>\>

The analysis highlighted an interesting finding, as shown in \@ref(fig:insight1): a significant concentration of COVID-19 cases appears within the 1500-2500 km range from the equator, suggesting a strong relationship between COVID-19 cases and cyclone activity within a margin of approximately 9 degrees of longitude. This area also highlights a geographical strip pointing to a genreal trend of lower GDP, government stability, and overall happiness.

```{r insight1, fig.cap=""}

covid_2020_world <- covid_2020 |>
  filter(
    total_cases < 50000000 
  )

ggplot(covid_2020_world, aes(x = distance_to_equator_km, y = total_cases)) +
  geom_point(color = "steelblue", alpha = 0.3) +  # Lower alpha for more transparency
  geom_density_2d(color = "darkorange") +  # Density lines to show clustering
  geom_vline(xintercept = 1500, linetype = "dashed", color = "black") +
  geom_vline(xintercept = 2500, linetype = "dashed", color = "black")  +
  labs(
    title = "Scatter Plot of COVID-19 Cases by Distance from Equator",
    x = "Distance from Equator (km)",
    y = "Total COVID-19 Cases"
  ) +
  coord_cartesian(xlim = c(0, 10250)) +
  scale_y_continuous(
    labels = scales::comma,
    trans = "log10",
    breaks = c(1, 10, 1e2, 1e3, 1e4, 1e5, 1e6, 1e7, 5e7)  
  ) +
  theme_minimal()

ggplot(cyclones_data |> filter(!is.na(category)), aes(x = km_distance_from_equator, y = category, fill = category)) +
  geom_density_ridges(scale = 2, alpha = 0.7, color = "black") +
  geom_vline(xintercept = 1500, linetype = "dashed", color = "black") +
  geom_vline(xintercept = 2500, linetype = "dashed", color = "black") +
  labs(
    x = "Distance from Equator (km)",
    y = "Cyclone Category",
    title = "Ridgeline Plot of Distance from Equator by Cyclone Category"
  ) +
  coord_cartesian(xlim = c(0, 6500)) +
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.grid.major = element_line(color = "grey80"),
    panel.grid.minor = element_blank(),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10)
  )

```

A report conducted by the [National Librarbty of Medicine](https://pmc.ncbi.nlm.nih.gov/articles/PMC8079387/) highlights a strong correlation between COVID-19 cases and proximity to the equator, suggesting that certain geographic and environmental factors may influence virus spread and outbreak. This observation signifies a trend where regions near the equator are subject to unique vulnerabilities, not only experiencing increased COVID-19 cases but also facing intensified cyclone activity (as shown in the figure above), volcanic activity, elevated ocean currents, seasonal flooding, and the presence of natural carbon sinks.

Applying this insight into the analysis, it is evident that there are strong correlations to geological crises such as cyclones and COVID-19, and key indicators like GDP, life satisfaction, social support, and the overall well-being of countries.These crises profoundly impact not only economic stability but also the social and health of a nation, as evidenced by happiness levels in Figure 5 of the appendix. This reinforces the insight that crisis have a strong effect on the affected regions economic and social dynamics.

```{r insight2, fig.cap=""}
#Social support & life ladder

data_combined <- inner_join(covid_2020, happiness, by = "country")

median_social_support <- median(data_combined$social_support, na.rm = TRUE)

ggplot(data_combined, aes(x = social_support, y = life_ladder, color = continent)) +
  geom_point(alpha = 0.7, position = "jitter") + 
  geom_smooth(method = "lm", color = "blue", se = TRUE) + 
  geom_vline(xintercept = median_social_support, color = "red", linetype = "dashed", linewidth = 1) + 
  labs(
    title = "Comparison of Social Support and Life Ladder by Continent",
    x = "Social Support",
    y = "Life Ladder (Happiness Score)",
    color = "Continent"
  ) +
  facet_wrap(~ continent) + 
  theme_minimal() +
  theme(panel.grid.major = element_line(color = "grey", linetype = "dashed", linewidth = 0.5))

# Social Support & Life Expectancy

ggplot(data_combined, aes(x = social_support, y = healthy_life_expectancy_at_birth, color = continent)) +
  geom_point(alpha = 0.7, position = "jitter") + 
  geom_smooth(method = "lm", color = "blue", se = TRUE) + 
  geom_vline(xintercept = median_social_support, color = "red", linetype = "dashed", linewidth = 1) + 
  labs(
    title = "Comparison of Social Support and Life Expectancy by Continent",
    x = "Social Support",
    y = "Healthy Life Expectancy at Birth",
    color = "Continent"
  ) +
  facet_wrap(~ continent) + 
  theme_minimal() +
  theme(panel.grid.major = element_line(color = "grey", linetype = "dashed", linewidth = 0.5))

data_combined <- inner_join(covid_2020, happiness, by = "country")

median_total_cases <- median(data_combined$total_cases, na.rm = TRUE)

ggplot(data_combined, aes(x = total_cases / 100, y = freedom_to_make_life_choices)) +
  geom_point(alpha = 0.5, position = "jitter") + 
  geom_smooth(method = "lm", color = "blue", se = TRUE) + 
  geom_vline(xintercept = median_total_cases, color = "red", linetype = "dashed", linewidth = 1) + 
  labs(
    title = "Comparison of COVID-19 Cases and Freedom",
    x = "Total COVID-19 Cases",
    y = "Freedom"
  ) +
  scale_x_log10() +  
  scale_y_log10() + 
  theme_minimal() +
  theme(panel.grid.major = element_line(color = "grey", linetype = "dashed", linewidth = 0.5))

#Life expectancy & Cyclone count

cyclone_counts <- cyclones_data |>
  group_by(country) |>
  summarise(cyclone_count = n())

combined_data <- cyclone_counts |>
  inner_join(happiness, by = "country") |>
  select(country, cyclone_count, healthy_life_expectancy_at_birth)

combined_data <- combined_data |>
  arrange(desc(cyclone_count)) |>
  slice(-1:-3)

ggplot(combined_data, aes(y = cyclone_count, x = healthy_life_expectancy_at_birth)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = TRUE, color = "blue") +
  labs(
    title = "Relationship Between Cyclone Count and Life Expectancy (Outlier Removed)",
    x = "Cyclone Count",
    y = "Healthy Life Expectancy at Birth"
  ) +
  theme_minimal()

colnames(happiness)

# distance & cyclone count

# Load required libraries
library(tidyverse)

# Step 1: Aggregate cyclone counts by country
cyclone_counts <- cyclones_data |>
  group_by(country) |>
  summarise(cyclone_count = n())

# Step 2: Join with happiness data on 'country' to add distance from the equator
combined_data <- cyclone_counts |>
  inner_join(happiness, by = "country") |>
  select(country, cyclone_count, distance_to_equator_km)

# Step 3: Remove the top 3 highest cyclone counts
combined_data <- combined_data |>
  arrange(desc(cyclone_count)) |>
  slice(-1:-3)

# Step 4: Plot the relationship between distance to equator and cyclone count
ggplot(combined_data, aes(x = distance_to_equator_km, y = cyclone_count)) +
  geom_point(alpha = 0.7, color = "blue") +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  labs(
    title = "Comparison of Cyclone Count and Distance from Equator (Top 3 Outliers Removed)",
    x = "Distance from Equator (km)",
    y = "Cyclone Count"
  ) +
  theme_minimal()

ggplot(cyclones_data |> filter(!is.na(category)), aes(x = km_distance_from_equator, y = category, fill = category)) +
  geom_density_ridges(scale = 2, alpha = 0.7, color = "black", size = 0.3) +
  labs(
    x = "Distance from Equator (km)",
    y = "Cyclone Category",
    title = "Ridgeline Plot of Distance from Equator by Cyclone Category"
  ) +
  coord_cartesian(xlim = c(0, 6500)) +
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.grid.major = element_line(color = "grey80"),
    panel.grid.minor = element_blank(),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10)
  )

colnames(happiness)

ggplot(cyclones_data |> filter(!is.na(category)), aes(x = km_distance_from_equator, y = category, fill = category)) +
  geom_density_ridges(scale = 2, alpha = 0.7, color = "black") +
  annotate("rect", xmin = 1500, xmax = 2500, ymin = 0, ymax = Inf, fill = "orange", alpha = 0.8) +
  labs(
    x = "Distance from Equator (km)",
    y = "Cyclone Category",
    title = "Ridgeline Plot of Distance from Equator by Cyclone Category"
  ) +
  coord_cartesian(xlim = c(0, 6500)) +
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.grid.major = element_line(color = "grey80"),
    panel.grid.minor = element_blank(),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10)
  )

happiness <- happiness|>
  left_join(
    rnaturalearth::ne_countries(scale = "medium", returnclass = "sf") |>
      transmute(
        country = name_long,
        distance_to_equator_km = geosphere::distGeo(
          cbind(st_coordinates(st_centroid(geometry))[,1], st_coordinates(st_centroid(geometry))[,2]),
          cbind(st_coordinates(st_centroid(geometry))[,1], 0)
        ) / 1000  # Convert to km
      ),
    by = "country"
  )

happiness

happiness_bands_gdp <- happiness |>
  mutate(distance_band = cut(distance_to_equator_km.y, breaks = seq(0, 10000, by = 500), include.lowest = TRUE, labels = seq(0, 9500, by = 500))) |>
  group_by(distance_band) |>
  summarize(avg_log_gdp_per_capita = mean(log_gdp_per_capita, na.rm = TRUE))

# Plot for log GDP per capita
ggplot(happiness_bands_gdp, aes(x = distance_band, y = avg_log_gdp_per_capita)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  annotate("rect", xmin = "1500", xmax = "2500", ymin = 0, ymax = Inf, fill = "orange", alpha = 0.8) +
  labs(
    title = "Average GDP per Capita by Distance from Equator",
    x = "Distance from Equator (km, Binned)",
    y = "Average Log GDP per Capita"
  ) +
  scale_x_discrete(breaks = seq(0, 9500, by = 1000)) +
  coord_cartesian(ylim = c(min(happiness_bands_gdp$avg_log_gdp_per_capita, na.rm = TRUE) - 0.1, 
                           max(happiness_bands_gdp$avg_log_gdp_per_capita, na.rm = TRUE) + 0.1)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5))

# Group data into 500 km distance bands and calculate average life ladder score for each band
happiness_bands_life <- happiness |>
  mutate(distance_band = cut(distance_to_equator_km.y, breaks = seq(0, 10000, by = 500), include.lowest = TRUE, labels = seq(0, 9500, by = 500))) |>
  group_by(distance_band) |>
  summarize(avg_life_ladder = mean(life_ladder, na.rm = TRUE))

# Plot for life ladder
ggplot(happiness_bands_life, aes(x = distance_band, y = avg_life_ladder)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  annotate("rect", xmin = "1500", xmax = "2500", ymin = 0, ymax = Inf, fill = "orange", alpha = 0.8) +
  labs(
    title = "Average Life Ladder Score by Distance from Equator",
    x = "Distance from Equator (km, Binned)",
    y = "Average Life Ladder Score"
  ) +
  scale_x_discrete(breaks = seq(0, 9500, by = 1000)) +
  coord_cartesian(ylim = c(min(happiness_bands_life$avg_life_ladder, na.rm = TRUE) - 0.1, 
                           max(happiness_bands_life$avg_life_ladder, na.rm = TRUE) + 0.1)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5))
```

This insight is supported by the summary statistics in table \@ref(tab:summary_stats)

The next insight that we found is shown in \@ref(fig:insight2).

```{r insight3, fig.height=4, fig.width=6, fig.cap="This is a figure caption that you will need to change in order to get good marks in the visualization rubric items."}

covid_happiness <- inner_join(covid_2020_world, happiness, by = "country")

filter_covid_happiness <- covid_happiness |> filter(!is.na(total_cases), !is.na(perceptions_of_corruption))

ratio_covid_happiness <- filter_covid_happiness |> group_by(country) |>
  mutate(
    ratio = total_cases * perceptions_of_corruption
  )

covid_happiness |> filter(!is.na(total_cases), !is.na(perceptions_of_corruption)) |>
  ggplot(mapping = aes(x = perceptions_of_corruption, y = total_cases, color = continent)) + geom_boxplot() + facet_wrap(~continent) + labs(title = "Perception of Corruption and COVID cases by continent", subtitle = "Weight of COVID cases repective to perception of corruption", x = "Perception of Corruption (0 - 1)", y = "Total Cases", caption = "Source: covid.ourworldindata.org\nworldhappiness.report") + theme_minimal() + theme(legend.position = "None")

```

Finally, \@ref(fig:insight3) shows ...

```{r insight3, fig.height=4, fig.width=6, fig.cap="This is a figure caption that you will need to change in order to get good marks in the visualization rubric items."}

```

# Conclusion and Future Work

Overall, we found \<<general ideas>\>.

A second paragraph about our findings.

The next steps in this analysis are...

The limitations of this analysis are as follows. (Do not simply list potential issues with sampling, but relate them to your analysis and how they affect your conclusions. An honest and complete acknowledgement of the limitations makes the analysis more trustworthy.)

# References

I am not strict about MLA or APA style or anything like that. For this report, I would much rather have your citations be easy to match to your insights.

The easiest way is to use Rmd's [footnote](https://bookdown.org/yihui/rmarkdown/markdown-syntax.html#inline-formatting) syntax. This will put a number beside the word where the footnote appears, and the full text of the footnote at the bottom of the page (pdf) or end of the document (html). The syntax is:[^1], where I suggest that you put in something like this[^2] to make references for this assignment.

[^1]: See the source view to see this footnote

[^2]: The relevance to the insight is ... . From \<<name of source and name of article>\>, published on \<<date>\>, url: \<<link to page>\>

Alternatively, you could make a list of citations with their main arguments and why they're relevent to your insights, methods, etc.

The link above also references "bibtex" files. These are also extremely convenient, but have a steep learning curve and they make it difficult to tie them to an insight. If you use bibtext, then make sure that you provide a sentence to describe the source and it's relevance when you cite it - don't just add citations to the end of a sentence (this is common practice in academia, but I want to know that your citations are directly relevant for this assignmnet).

# Apendix

```{r world_map_cyclones_by_category, fig.cap="North Atlantic and Northeast Pacific Cyclones by Category"}
world_map <- map_data("world")

cyclones_data <- cyclones_data |>
    mutate(MonthName = factor(month.abb[Month], levels = month.abb)) |>
    relocate(MonthName, .after = Month)

ggplot() +
    geom_polygon(data = world_map, aes(x = long, y = lat, group = group), fill = "lightgray", color = "white") +
    geom_rect(aes(xmin = -180, xmax = 180, ymin = -5, ymax = 5), fill = "red", alpha = 0.2) + 
    geom_hex(data = cyclones_data, aes(x = long, y = lat, fill = category), bins = 50, alpha = 0.7) +
    scale_fill_viridis_d(name = "Storm Category") +
    labs(
        title = "North Atlantic and Northeast Pacific Cyclones by Category",
        x = "Longitude",
        y = "Latitude"
    ) +
    scale_x_continuous(breaks = c(-180, -90, 0, 90, 180)) +
    scale_y_continuous(breaks = c(-90, -45, 0, 45, 90)) +
    theme_minimal()
```

```{r world_map_cyclones_by_basin, fig.cap="North Atlantic and Northeast Pacific Cyclones by Basin"}
ggplot() +
    geom_polygon(data = world_map, aes(x = long, y = lat, group = group), fill = "lightgray", color = "white") +
    geom_rect(aes(xmin = -180, xmax = 180, ymin = -5, ymax = 5), fill = "red", alpha = 0.2) +
    geom_hex(data = cyclones_data, aes(x = long, y = lat, fill = Basin, alpha = ..count..), bins = 50) +
    scale_fill_viridis_d(name = "Storm Category") +
    scale_alpha(range = c(0.35, 1.5)) +
    labs(
        title = "North Atlantic and Northeast Pacific Cyclones by Basin",
        x = "Longitude",
        y = "Latitude"
    ) +
    scale_x_continuous(breaks = c(-180, -90, 0, 90, 180)) +
    scale_y_continuous(breaks = c(-90, -45, 0, 45, 90)) +
    theme_minimal()
```

```{r storm_frequency_by_month_year, fig.cap="Storm Frequency by Month and Year"}
ggplot(cyclones_data) +
    aes(x = ObservYear, y = MonthName, fill = ..count..) +
    geom_bin2d() +
    scale_fill_distiller(direction = 1, palette = "YlOrRd") +
    labs(
        title = "Storm Frequency by Month and Year",
        x = "Observation Year",
        y = "Month",
        fill = "Frequency"
    ) +
    theme_minimal()
```

```{r proportion_storm_categories_over_time, fig.cap="Proportion of Storm Categories Over Time"}
ggplot(cyclones_data) +
    aes(x = ObservYear, fill = category) +
    geom_bar(position = "fill", width = 1) +
    labs(
        title = "Proportion of Storm Categories Over Time",
        x = "Observation Year",
        y = "Proportion",
        fill = "Storm Category"
    ) +
    theme_minimal()
```

```{r world_happiness_scores, fig.cap="World Happiness Scores by Country"}
happiness_colors <- c("1" = "#000000", "3" = "#aa0000", "4" = "#f46d43",
                      "5" = "#fdae61", "6" = "#fee08b", "7" = "#66bd63",
                      "8" = "#1a9850")

ggplot(happiness_mapped) +
    aes(x = long, y = lat, group = group, fill = factor(round(life_ladder))) +
    geom_polygon(color = "white") +
    scale_fill_manual(values = happiness_colors, name = "Happiness\nScore") +
    theme_void() +
    labs(title = "World Happiness Scores by Country")
```