### 2024-11-15 Draft.Rmd
brushed

---
title: "Analysis of Global Crises and Their Impact on Interational Well-Being"
author: "Group 8"
date: "November 12, 2024"
output: pdf_document
---

List your group members, including their student numbers, here:

- Aditya Chauhan (169027493)
- Andy Ferrer Lorenzo (169081356)
- Oscar Saul (169057882)
- Taras Zakydalsky (169099616)
- Dibuhang Rai (169708451)

You **must** be in a group in MyLS in order to see the DropBox used for submission. Even if you're alone, you must join a group by yourself.

You **must** be in a group with people from the same section as you. MyLS does not allow for groups including students from both Data100A and Data100B.

```{r initial_setup, include=FALSE}
# echo = FALSE will set the Rmd to *not* show the R code. Don't change this.
# You may change the default figure width and figure height as you please.
knitr::opts_chunk$set(echo = FALSE, message = FALSE, fig.width = 6)
# Put any libraries that you need to load here.
# DO NOT PUT "install.packages()" IN AN RMD FILE!!!
library(tidyverse)
library(arrow)
library(maps)
library(tinytex)
library(sf)
library(rnaturalearth)
library(corrplot)
library(scales)
library(geosphere)
library(ggridges)
```

# Purpose
This study aims to investigate the relationship between freedom of citizens (i.e. freedom of choices, a just government) and quality of life for citizens. The more freedom for citizens allows for better quality of life.

# Abstract

This paper will aim to show as freedom of citizens increase and citizen's quality of life improves. The relationship we look for is a positive correlation. First, for the first insight, we explore the relationship of cyclone frequencies and geological location. Cyclones of higher category appear more frequently near the equator (~2000 km from the equator), more freedom allows for an individual to choose where they can locate themselves and those who choose to live farther away from the equator may feel less burdened by constant cyclones. Second, a more just government allows for a citizen to live with more freedom, from th second insight, "perception of corruption" for a government and "COVID cases" shows us for some continents, there are lower cases of COVID when the government is perceived to be fair. Though there also exists a negative correlation, where continents (i.e. Asia) with more corrupt ruling powers have lower total count of COVID cases. Third, again back to the notion of choosing where to live, from the third insight, we observe average GDP in relation to distance from the equator. Areas farther from the equator tend to have higher on average higher GDP than those closer to the equator. We now show evidence to further support our original purpose. That freedom of citizens in fact impacts the quality of life for citizens.

# Introduction

Quality of life determines happiness which is impacted by many factors. Some important factors are environmental and socioeconomic factors. Environmentally we must keep in mind location, since different countries in different continents vary in climate, geographic identity, etc. Whilst socioeconomically, different countries have different forms of government, culture.

In this report, we will explore the geographical aspect and governance aspect of freedom and what do these factors impact on an individual's daily life through exploring and digesting our data. We will explore data; COVID cases during the year 2020, cyclones from NOAA (National Oceanic and Atmospheric Administration), happiness data from (World Happiness Report). We apply uni-variate and bi-variate analysis, specifically for the bard chart and ridge line graph we care about the frequency (weight) of our independent variables, for the scatter plot to determine the relationship between our variable x and y is applying bivariate analysis.

By the end of the report, we will have shown in general that as freedom increases for citizens and individuals, so to does quality of life.

# Data Description

## \<\<Cyclones Data\>\>

```{r cyclones_data}
cyclones_data <- read_parquet("cyclones_data.parquet")
world <- ne_countries(scale = "medium", returnclass = "sf")

# ________________Map long and lat to scale________________
# Convert latitude(-90/+90) and longitude(-180/+180) to be within the correct range
cyclones_data <- cyclones_data |>
    rename(long = lon) |> #rename lon to long to match with world map
    mutate(
        lat = ((lat + 90) %% 180) - 90,  
        long = ((long + 180) %% 360) - 180 
    )

# ________________Determine distance of cyclone to equator________________
equator_point <- c(0, 0)
cyclones_data <- cyclones_data |>
  mutate(
    km_distance_from_equator = distGeo(cbind(long, lat), cbind(long, 0)) / 1000
  )

# ________________Determine distance of cyclone to equator________________
equator_point <- c(0, 0)
cyclones_data <- cyclones_data |>
  mutate(
    km_distance_from_equator = distGeo(cbind(long, lat), cbind(long, 0)) / 1000
  )

# ________________Convert long and lat points to countries________________
cyclones_data <- st_as_sf(cyclones_data, coords = c("long", "lat"), crs = 4326, remove = FALSE) |>
  st_join(world["name"], left = TRUE) |>
  rename(country = name)

# ________________Create cyclone_cleaned with specified columns________________
cyclone_cleaned <- cyclones_data |>
    select(country, lat, long, category, km_distance_from_equator)

cyclone_cleaned

```

The data set `cyclone_data` provided by the National Hurricane Center and the Central Pacific Hurricane Center records hurricane entries in the Atlantic and North Pacific basins from 1851 to 2022 and 1949 to 2022, respectively. The data frame includes detailed time stamps, coordinates, and categories, which are vital for analysis.

The preliminary step in supporting the analysis involved converting the coordinates into the Cartesian coordinate system. Next, we used the `rnaturalearth` package and `geolocation` to convert the coordinates into country identifiers and determine each storm's distance in kilometers from the equator. As shown in Figure 1 in the appendix, cyclones are now accurately distributed across the North Atlantic and Northeast Pacific basins with the correct identifiers.

## \<\<World Covid 2020\>\>

```{r load_data2}
covid_2020 <- read_parquet("covid_2020.parquet")
world <- ne_countries(scale = "medium", returnclass = "sf")

covid_2020 <- covid_2020 |>
  mutate(country = case_when(
  country == "British Virgin Islands" ~ "British Virgin Is.",
  country == "Cape Verde" ~ "Cabo Verde",
  country == "Cayman Islands" ~ "Cayman Is.",
  country == "Central African Republic" ~ "Central African Rep.",
  country == "Cook Islands" ~ "Cook Is.",
  country == "Cote d'Ivoire" ~ "Ivory Coast",
  country == "Democratic Republic of Congo" ~ "Congo",
  country == "Cyprus" ~ "Cyprus",
  country == "Dominican Republic" ~ "Dominican Rep.",
  country == "East Timor" ~ "Timor-Leste",
  country == "England" ~ "United Kingdom",
  country == "Equatorial Guinea" ~ "Equat. Guinea",
  country == "Eswatini" ~ "Swaziland",
  country == "Falkland Islands" ~ "Falkland Isles",
  country == "Faroe Islands" ~ "Faroe Isles",
  country == "French Guiana" ~ "Fr. Guiana",
  country == "French Polynesia" ~ "Fr. Polynesia",
  country == "Gibraltar" ~ "Gibraltar",
  country == "Guadeloupe" ~ "Guadeloupe",
  country == "Marshall Islands" ~ "Marshall Is.",
  country == "Martinique" ~ "Martinique",
  country == "Mayotte" ~ "Mayotte",
  country == "Micronesia (country)" ~ "Micronesia",
  country == "Northern Ireland" ~ "Northern Ireland",
  country == "Northern Mariana Islands" ~ "N. Mariana Is.",
  country == "Pitcairn Islands" ~ "Pitcairn",
  country == "Reunion" ~ "Réunion",
  country == "Saint Barthelemy" ~ "Saint-Barthélemy",
  country == "Saint Kitts and Nevis" ~ "St Kitts & Nevis",
  country == "Saint Martin (French part)" ~ "Saint Martin",
  country == "Saint Pierre and Miquelon" ~ "St Pierre & Miquelon",
  country == "Saint Vincent and the Grenadines" ~ "St Vincent & Gren.",
  country == "Sao Tome and Principe" ~ "São Tomé",
  country == "Scotland" ~ "Scotland",
  country == "Sint Maarten (Dutch part)" ~ "Sint Maarten",
  country == "Solomon Islands" ~ "Solomon Is.",
  country == "South Sudan" ~ "S. Sudan",
  country == "Tokelau" ~ "Tokelau",
  country == "Turks and Caicos Islands" ~ "Turks & Caicos",
  country == "United States Virgin Islands" ~ "US Virgin Is.",
  country == "Wales" ~ "Wales",
  country == "Wallis and Futuna" ~ "Wallis & Futuna",
  country == "United States" ~ "United States of America",
  TRUE ~ country  
))

equator <- st_sfc(st_linestring(rbind(c(-180, 0), c(180, 0))), crs = st_crs(world))
world <- st_transform(world, crs = 3857)
equator <- st_transform(equator, crs = 3857)

world_with_distance <- world |>
  mutate(
    centroid = st_centroid(geometry),
    latitude = st_coordinates(centroid)[, 2],
    hemisphere = ifelse(latitude > 0, "North", "South"),
    distance_to_equator_km = as.numeric(st_distance(centroid, equator)) / 1000  # Convert to kilometers
  ) |>
  select(name, hemisphere, distance_to_equator_km) |>
  rename(country = name)

covid_2020 <- covid_2020 |>
  left_join(world_with_distance, by = "country")

covid_2020

```

The `covid_2020` data set, provided by Our World in Data contains a comprehensive list of the total reported cases of COVID-19 cases across many countries and continents in 2020. To aid in the analysis `rnaturalearth` and `geolocation` was used to find the hemisphere and distance to the equator in kilometers.

## \<\<World Happiness Census\>\>

```{r load_data3}
# Put in your code to load in the data set, along with any
# necessary cleaning beyond what was done in Part 1
happiness <- read_parquet("happiness.parquet")
world_map <- map_data("world")

happiness_mapped <- happiness |>
    mutate(country = case_when( #match country names to world_map
        country == "United States" ~ "United States of America",
        country == "United Kingdom" ~ "UK",
        country == "Congo (Brazzaville)" ~ "Congo",
        country == "Congo (Kinshasa)" ~ "Congo",
        country == "Hong Kong S.A.R. of China" ~ "Hong Kong",
        country == "Türkiye" ~ "Turkey",
        country == "Lao.People's.Democratic.Republic" ~ "Laos",
        country == "United Arab Emirates" ~ "UAE",
        TRUE ~ country
    )) |>
    right_join(world_map, by = c("country" = "region")) #joins the long and lat of the respective countries to map them

world <- ne_countries(scale = "medium", returnclass = "sf")

equator <- st_sfc(st_linestring(rbind(c(-180, 0), c(180, 0))), crs = st_crs(world))
world <- st_transform(world, crs = 3857)
equator <- st_transform(equator, crs = 3857)

happiness <- happiness |>
  left_join(
    world |>
      mutate(
        centroid = st_centroid(geometry),
        latitude = st_coordinates(centroid)[, 2],
        hemisphere = ifelse(latitude > 0, "North", "South"),
        distance_to_equator_km = as.numeric(st_distance(centroid, equator)) / 1000
      ) |>
      select(name, hemisphere, distance_to_equator_km) |>
      rename(country = name),
    by = "country"
  )

```

The `happiness` data set is global census on a countries economic, social, and political stability provided by World Happiness Report. From part 1, extra steps were applied to clean the data set. Some steps taken are renaming countries to allow for easy joining of data sets (happiness dataset and world map). For this case only specific columns were needed and hence selected specific columns.

## Combining the Data

For the first data set, Cyclones data "sf_join" was used to create geographical data for a map for distances of cyclones from the equator (this was not used but is in the appendix). Only necessary columns were selected from the data set.
For the second data set, World data was used from 'rnaturalearth' package. First countries were renamed to allow for easier joining of data sets. New columns were created for the world map to be created. Then the data set is combined with COVID_2020 by country.
For the third data set, country names were changed for easier merge. Then applying "right_join" to keep the data from world_map and combine with happiness report data. This again is combined with world data, this time applying "left_join" to keep the data from happiness data and combine with world data.

# Exploratory Data Analysis

We explored many aspects of the data, but will demonstrate three. These are \<\<Geographic and environmental Factors VS a countries overall well being\>\>, \<\<Covid Cases Vs perception of corruption\>\>, and \<<Life ladder and GDP vs Distance from Equator>\>

The analysis highlighted an interesting finding, as shown in \@ref(fig:insight1): a significant concentration of COVID-19 cases appears within the 1500-2500 km range from the equator, suggesting a strong relationship between COVID-19 cases and cyclone activity within a margin of approximately 9 degrees of longitude. This area also highlights a geographical strip pointing to a general trend of lower GDP, government stability, and overall happiness.

```{r insight1, fig.cap=""}

ggplot(cyclones_data |> filter(!is.na(category)), aes(x = km_distance_from_equator, y = category, fill = category)) +
  geom_density_ridges(scale = 2, alpha = 0.7, color = "black") +
  geom_vline(xintercept = 1500, linetype = "dashed", color = "black") +
  geom_vline(xintercept = 2500, linetype = "dashed", color = "black") +
  labs(
    x = "Distance from Equator (km)",
    y = "Cyclone Category",
    title = "Ridgeline Plot of Distance from Equator by Cyclone Category"
  ) +
  coord_cartesian(xlim = c(0, 6500)) +
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.grid.major = element_line(color = "grey80"),
    panel.grid.minor = element_blank(),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10)
  )

```

A report conducted by the [National Library of Medicine](https://pmc.ncbi.nlm.nih.gov/articles/PMC8079387/) highlights a strong correlation between COVID-19 cases and proximity to the equator, suggesting that certain geographic and environmental factors may influence virus spread and outbreak. This observation signifies a trend where regions near the equator are subject to unique vulnerabilities, not only experiencing increased COVID-19 cases but also facing intensified cyclone activity (as shown in the figure above), volcanic activity, elevated ocean currents, seasonal flooding, and the presence of natural carbon sinks.

Applying this insight into the analysis, it is evident that there are strong correlations to geological crises such as cyclones and COVID-19, and key indicators like GDP, life satisfaction, social support, and the overall well-being of countries.These crises profoundly impact not only economic stability but also the social and health of a nation, as evidenced by happiness levels in Figure 5 of the appendix. This reinforces the insight that crisis have a strong effect on the affected regions economic and social dynamics.

The next insight that we found is shown in \@ref(fig:insight2).

```{r insight2, fig.height=4, fig.width=6, fig.cap=""}

covid_happiness <- inner_join(covid_2020, happiness, by = "country")

filter_covid_happiness <- covid_happiness |> filter(!is.na(total_cases), !is.na(perceptions_of_corruption))

ratio_covid_happiness <- filter_covid_happiness |> group_by(country) |>
  mutate(
    ratio = total_cases * perceptions_of_corruption
  )

covid_happiness |> filter(!is.na(total_cases), !is.na(perceptions_of_corruption), total_cases <= 20000000) |>
  ggplot(mapping = aes(x = perceptions_of_corruption, y = total_cases, color = continent)) + geom_point() + geom_smooth(method = "lm", se = FALSE) + facet_wrap(~continent) + labs(title = "Perception of Corruption and COVID cases by continent", subtitle = "Weight of COVID cases respective to perception of corruption", x = "Perception of Corruption (0 - 1)", y = "Total Cases", caption = "Source: covid.ourworldindata.org\nworldhappiness.report") + theme_minimal() + theme(legend.position = "None")

```

This insight is supported by the summary statistics in table \@ref(tab:summary_stats)

```{r summary_stats, fig.height=4, fig.width=6,}
library(gtsummary)
library(gapminder)

insight_2_sum_stats <- covid_happiness |> group_by(continent) |> select(continent, perceptions_of_corruption, total_cases) |> 
  tbl_summary(by = continent)
insight_2_sum_stats
```

From [Scientific Reports (Sci Rep)](https://www.nature.com/articles/s41598-021-02802-1), the paper implies that corruption leads to lower rates of vaccination. This can result in higher counts of COVID cases. Also [Globalization and Health](https://globalizationandhealth.biomedcentral.com/articles/10.1186/s12992-023-00994-x) also accounts to how corruption can manifest due to catastrophic events such as a pandemic. Then corruption affects how we tackle on health issues and health issues also affect how corruption is procured in the globlal stage.

Finally, \@ref(fig:insight3) shows ...

```{r insight3, fig.height=4, fig.width=6, fig.cap=""}
happiness <- happiness|>
  left_join(
    rnaturalearth::ne_countries(scale = "medium", returnclass = "sf") |>
      transmute(
        country = name_long,
        distance_to_equator_km = geosphere::distGeo(
          cbind(st_coordinates(st_centroid(geometry))[,1], st_coordinates(st_centroid(geometry))[,2]),
          cbind(st_coordinates(st_centroid(geometry))[,1], 0)
        ) / 1000  # Convert to km
      ),
    by = "country"
  )

happiness_bands_gdp <- happiness |>
  mutate(distance_band = cut(distance_to_equator_km.y, breaks = seq(0, 10000, by = 500), include.lowest = TRUE, labels = seq(0, 9500, by = 500))) |>
  group_by(distance_band) |>
  summarize(avg_log_gdp_per_capita = mean(log_gdp_per_capita, na.rm = TRUE))

# Plot for log GDP per capita
ggplot(happiness_bands_gdp, aes(x = distance_band, y = avg_log_gdp_per_capita)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  annotate("rect", xmin = "1500", xmax = "2500", ymin = 0, ymax = Inf, fill = "orange", alpha = 0.8) +
  labs(
    title = "Average GDP per Capita by Distance from Equator",
    x = "Distance from Equator (km, Binned)",
    y = "Average Log GDP per Capita"
  ) +
  scale_x_discrete(breaks = seq(0, 9500, by = 1000)) +
  coord_cartesian(ylim = c(min(happiness_bands_gdp$avg_log_gdp_per_capita, na.rm = TRUE) - 0.1, 
                           max(happiness_bands_gdp$avg_log_gdp_per_capita, na.rm = TRUE) + 0.1)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5))

```

# Conclusion and Future Work

Overall, we found \<<general ideas>\>.

A second paragraph about our findings.

The next steps in this analysis are...

The limitations of this analysis are as follows. (Do not simply list potential issues with sampling, but relate them to your analysis and how they affect your conclusions. An honest and complete acknowledgement of the limitations makes the analysis more trustworthy.)

# References

I am not strict about MLA or APA style or anything like that. For this report, I would much rather have your citations be easy to match to your insights.

The easiest way is to use Rmd's [footnote](https://bookdown.org/yihui/rmarkdown/markdown-syntax.html#inline-formatting) syntax. This will put a number beside the word where the footnote appears, and the full text of the footnote at the bottom of the page (pdf) or end of the document (html). The syntax is:[^1], where I suggest that you put in something like this[^2] to make references for this assignment.

[^1]: See the source view to see this footnote

[^2]: The relevance to the insight is ... . From \<<name of source and name of article>\>, published on \<<date>\>, url: \<<link to page>\>

Alternatively, you could make a list of citations with their main arguments and why they're relevent to your insights, methods, etc.

The link above also references "bibtex" files. These are also extremely convenient, but have a steep learning curve and they make it difficult to tie them to an insight. If you use bibtext, then make sure that you provide a sentence to describe the source and it's relevance when you cite it - don't just add citations to the end of a sentence (this is common practice in academia, but I want to know that your citations are directly relevant for this assignmnet).

# Apendix

```{r world_map_cyclones_by_category, fig.cap="North Atlantic and Northeast Pacific Cyclones by Category"}
world_map <- map_data("world")

cyclones_data <- cyclones_data |>
    mutate(MonthName = factor(month.abb[Month], levels = month.abb)) |>
    relocate(MonthName, .after = Month)

ggplot() +
    geom_polygon(data = world_map, aes(x = long, y = lat, group = group), fill = "lightgray", color = "white") +
    geom_rect(aes(xmin = -180, xmax = 180, ymin = -5, ymax = 5), fill = "red", alpha = 0.2) + 
    geom_hex(data = cyclones_data, aes(x = long, y = lat, fill = category), bins = 50, alpha = 0.7) +
    scale_fill_viridis_d(name = "Storm Category") +
    labs(
        title = "North Atlantic and Northeast Pacific Cyclones by Category",
        x = "Longitude",
        y = "Latitude"
    ) +
    scale_x_continuous(breaks = c(-180, -90, 0, 90, 180)) +
    scale_y_continuous(breaks = c(-90, -45, 0, 45, 90)) +
    theme_minimal()
```

```{r world_map_cyclones_by_basin, fig.cap="North Atlantic and Northeast Pacific Cyclones by Basin"}
ggplot() +
    geom_polygon(data = world_map, aes(x = long, y = lat, group = group), fill = "lightgray", color = "white") +
    geom_rect(aes(xmin = -180, xmax = 180, ymin = -5, ymax = 5), fill = "red", alpha = 0.2) +
    geom_hex(data = cyclones_data, aes(x = long, y = lat, fill = Basin, alpha = ..count..), bins = 50) +
    scale_fill_viridis_d(name = "Storm Category") +
    scale_alpha(range = c(0.35, 1.5)) +
    labs(
        title = "North Atlantic and Northeast Pacific Cyclones by Basin",
        x = "Longitude",
        y = "Latitude"
    ) +
    scale_x_continuous(breaks = c(-180, -90, 0, 90, 180)) +
    scale_y_continuous(breaks = c(-90, -45, 0, 45, 90)) +
    theme_minimal()
```

```{r storm_frequency_by_month_year, fig.cap="Storm Frequency by Month and Year"}
ggplot(cyclones_data) +
    aes(x = ObservYear, y = MonthName, fill = ..count..) +
    geom_bin2d() +
    scale_fill_distiller(direction = 1, palette = "YlOrRd") +
    labs(
        title = "Storm Frequency by Month and Year",
        x = "Observation Year",
        y = "Month",
        fill = "Frequency"
    ) +
    theme_minimal()
```

```{r proportion_storm_categories_over_time, fig.cap="Proportion of Storm Categories Over Time"}
ggplot(cyclones_data) +
    aes(x = ObservYear, fill = category) +
    geom_bar(position = "fill", width = 1) +
    labs(
        title = "Proportion of Storm Categories Over Time",
        x = "Observation Year",
        y = "Proportion",
        fill = "Storm Category"
    ) +
    theme_minimal()
```

```{r world_happiness_scores, fig.cap="World Happiness Scores by Country"}
happiness_colors <- c("1" = "#000000", "3" = "#aa0000", "4" = "#f46d43",
                      "5" = "#fdae61", "6" = "#fee08b", "7" = "#66bd63",
                      "8" = "#1a9850")

ggplot(happiness_mapped) +
    aes(x = long, y = lat, group = group, fill = factor(round(life_ladder))) +
    geom_polygon(color = "white") +
    scale_fill_manual(values = happiness_colors, name = "Happiness\nScore") +
    theme_void() +
    labs(title = "World Happiness Scores by Country")
```